---
title: "NYPD_Data_Analysis"
author: "Anonymous"
output: pdf_document
---
```{r imports}
#import packages needed
library(tidyverse)
```

```{r import_nypd_data, message=FALSE}
#Assign a variable for the url of the csv data and then loads it
url <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
nypd_data <- read_csv(url)
```

```{r tidy_nypd_data}
#remove the columns corresponding to location descriptions 
nypd_data <- nypd_data %>% select(-c(LOC_OF_OCCUR_DESC, LOC_CLASSFCTN_DESC, 
                                     LOCATION_DESC, Latitude, Longitude)) 

#remove perpetrator columns as not all crimes have a perp description
nypd_data <- nypd_data %>% select(-c(PERP_AGE_GROUP, PERP_SEX, PERP_RACE))

#remove extra columns that are not needed
nypd_data <- nypd_data %>% select(-c(JURISDICTION_CODE, PRECINCT, INCIDENT_KEY, 
                                     X_COORD_CD, Y_COORD_CD, Lon_Lat, OCCUR_TIME, 
                                     STATISTICAL_MURDER_FLAG))
```

```{r transform_nypd_data}
#convert the full date string to a datetime object
nypd_data <- nypd_data %>% mutate(date = mdy(OCCUR_DATE)) %>% select(-c(OCCUR_DATE))

#final check that the character columns have non null data
nypd_data <- nypd_data %>% filter(!is.na(VIC_AGE_GROUP)) %>% 
  filter(!is.na(VIC_SEX)) %>% filter(!is.na(VIC_RACE)) %>% 
  filter(!is.na(BORO))

#summary of transformed nypd data
summary(nypd_data)
```

```{r visualize_nypd_data_by_race, message=FALSE}
#transform the data by creating a column for categories of victim's 
#race and the NYC borough where the crime occurred
crime_by_race_and_borough <- nypd_data %>%  filter(VIC_RACE!="UNKNOWN") %>% 
  group_by(BORO, VIC_RACE) %>% summarize(crimes = n_distinct(date)) %>%
  unite(race_borough,VIC_RACE,BORO,sep = "_")

#use a bar graph to visualize the data with the y axis being the categories of 
#race and borough and the x axis is the amount of crimes that occurred to that 
#category within the time period observed
crime_by_race_and_borough %>% ggplot(aes(x = crimes, y = race_borough)) + 
  geom_col(aes(color = "race_borough")) + 
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90)) + 
  labs(title = "Crimes in NYC Boroughs by Victim's Race",x="Number of Crimes", 
       y="Race and Borough")
```

```{r visualize_nypd_data_by_age_and_sex}
#transform the data by creating a column for categories of victim's age and sex 
#and then grouping the data by the new column
crime_by_age_sex = nypd_data %>% filter(VIC_AGE_GROUP!="UNKNOWN") %>% 
  filter(VIC_AGE_GROUP!="1022") %>% 
  filter(VIC_SEX!="U") %>% 
  unite(age_sex,VIC_AGE_GROUP,VIC_SEX,sep = "/") %>% 
  group_by(age_sex) %>% 
  summarize(crimes = n_distinct(date))

#use a bar graph to visualize the data with the y axis being the categories of 
#age and sex of the victims and the x axis is the amount of crimes that occurred 
#to that category within the time period observed
crime_by_age_sex %>% ggplot(aes(x = crimes, y = age_sex)) + 
  geom_col(aes(color = "age_sex")) + theme(legend.position = "bottom", 
                                           axis.text.x = element_text(angle = 90)) + 
  labs(title = "Crimes in NYC by Age and Sex of Victims", x="Number of Crimes", 
       y="Victim Age and Sex")
```

Analyzing the data of the first graph, I then took out the Asian and American Indian races as there isn't a lot of data recorded. I then eliminated the Black Hispanic category as there might be a co-linearity with the White Hispanic category. I then took the remaining races and took the borough with the highest count of crimes.

```{r visualize_borough_data_by_line_plot,message=FALSE}
crimes_by_date <- nypd_data %>%  filter(VIC_RACE!="UNKNOWN") %>% 
  unite(race_borough,VIC_RACE,BORO,sep = "_") %>% 
  filter(race_borough=="BLACK_BROOKLYN" | race_borough=="WHITE_BROOKLYN" | 
           race_borough=="WHITE HISPANIC_BRONX") %>% 
  group_by(race_borough,date) %>% summarize(crimes = n()) 

crimes_by_date %>% filter(crimes>0) %>% ggplot(aes(x = date, y = crimes)) + 
  geom_line(aes(color = race_borough)) + geom_point(aes(color = race_borough))+ 
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90)) + 
  labs(title = "Crimes in NYC Boroughs by Victim's Race",y="Number of Crimes", x="Date")
```

I then modeled the number of crimes as a function of the race_borough categories with 0 corresponding to the 'White Hispanic/Bronx' category, 1 corresponding to the 'Black Brooklyn' category and 2 corresponding to the 'White Brooklyn category'

```{r model_crime_borough_race_date}
nypd_data_model <- lm(as.numeric(ordered(race_borough))~crimes+
                        as.numeric(as.factor(date)),data = crimes_by_date)
summary(nypd_data_model)
crimes_with_prediction <- crimes_by_date %>% ungroup() %>% 
  mutate(pred = as.factor(round(predict(nypd_data_model))))

crimes_with_prediction %>% ggplot(aes(x = date, y = crimes)) + 
  geom_line(aes(color = race_borough)) + geom_point(aes(color = race_borough))+
  geom_line(aes(color = pred)) + geom_point(aes(color = pred))+
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90)) + 
  labs(title = "Crimes in NYC Boroughs by Victim's Race with Model Predictions",y="Number of Crimes", x="Date")
```
In conclusion, the model defaults to one class for the most part because a linear regression model can't capture the noise in the data. In terms of bias the data will tend towards having more incidents with black people as the source is the NYPD. I tried to point the data towards victims of the crime versus the perpetrator since I have a personal bias to care more about the victims. Another personal source of bias is that not every borough is equal and I am more familiar with Manhattan and Brooklyn so I tried to spread the data to more total categories to try to mitigate the bias.  
